Cross Validation Approaches for Penalized Cox Regression

### 2019-03-06 main.tex
clean/reply to some previous comments

\note{What's with all the \textbackslash par commands?  They don't seem to do anything.}

@@

\par In this paper, we propose two alternative ways to carry out cross validation for Cox regression. Instead of applying cross-validation to the partial likelihood directly, we propose applying cross validation to either the linear predictors of the regression model or to the deviance residuals \citep{Therneau1990}. Through simulation studies, we compare these proposed methods with existing cross-validation approaches for LASSO penalized Cox regression in both low- and high-dimensional settings.  \note{STILL TRUE?? We find that all cross validation approaches tend to be conservative,} \bnote{We should discuss this and the sentence in abstract at our next meeting. They are conservative only in terms of selecting the model that has the best preditive performance; they do not miss selecting variables in the true support}but that the linear predictor approach offers the best combination of performance and stability, and we recommend using it for regularization parameter selection in penalized Cox models.  We conclude by applying both proposed and existing approaches to two high dimensional data sets from real studies with time-to-event outcomes.

@@

The fundamental challenge of conducting cross validation for Cox regression is that the baseline hazard is not estimated from the model. Hence, another alternative approach to cross-validation would be to obtain an estimate $\hat{\Lambda}_{0}$ for the baseline hazard and use this in cross-validation.  There are several potential approaches to estimating $\Lambda_0$; in this manuscript, we chose to obtain the cross-validated linear predictors $\hat{\eta}^{cv}$ as in \eqref{eq:cv-lp}, then estimate $\Lambda_0$ via Kalbfleisch and Prentice's method \citep{Kalbfleisch2011}, treating these linear predictors as fixed inputs.\note{Maybe this isn't critical, but I still think we could just use $\hat{\Lambda}_0$ from the full data fit \ldots that estimator just makes more sense to me}\bnote{Biyue is taking a look at the baseline now. Discuss this at next meeting}

@@

Simulation studies were conducted to compare how those methods behave relative to each other in selecting the tuning parameter $\lambda$ and the corresponding model for LASSO penalized regression. Survival times were generated from exponential distribution $h(t) = h_{0} \exp(X\beta)$, where $h_{0}$ is the prespecified baseline\note{huh?  so what is $h_0$?}; the covariate matrix $X$ with $n$ observations and $p$ covariates is generated from independent $\text{Normal}(0, 1)$ samples; and the coefficient vector $\beta$ is sparse, with most entries equal to zero. Censoring status indicators were drawn independently from the binomial distribution. All simulations were implemented in R \citep{R}. We varied the sample size, dimension, censoring percentage and signal strength of the simulated data sets to examine how these factors affect the performance of cross-validation.

@@

Estimation accuracy was measured using the mean squared error of the coefficients ($\MSE$). Suppose $\hat{\beta}(\lambda)$ is the set of coefficients estimated by a cox model with LASSO penalty at $\lambda$. Mean squared error $MSE$ is defined to be 
\begin{equation}
E(\hat{\beta}(\lambda) - \beta)^2.
\end{equation}
\note{this is the equation for a scalar MSE, not a vector}

@@

As in Figure~\ref{Fig:mse-brier-c}, the standard and linear predictor approaches outperform the deviance residual and $V\&VH$ approaches.  The improvement is particularly noticeable in the most difficult setting: high-dimensional data with weak signal.  Table~\ref{Tab:sim} also provides some insight into why the deviance residual and $V\&VH$ approaches perform poorly relative to standard CV and the linear predictor approach: they act conservatively, consistently choosing larger $\lam$ values and selecting fewer variables. \note{if we're going to say something about ``selecting fewer variables'', should Table 1 include the average model size?}

@@

\begin{figure}[h]
    \centering
		\includegraphics[height= 15cm ]{./figures/figure_auc_extended.png}
    \caption{Cross-Validated AUC}
\end{figure}

>>>>

%\note{What's with all the \textbackslash par commands?  They don't seem to do anything.}

@@

\par In this paper, we propose two alternative ways to carry out cross validation for Cox regression. Instead of applying cross-validation to the partial likelihood directly, we propose applying cross validation to either the linear predictors of the regression model or to the deviance residuals \citep{Therneau1990}. Through simulation studies, we compare these proposed methods with existing cross-validation approaches for LASSO penalized Cox regression in both low- and high-dimensional settings.  %\note{STILL TRUE?? We find that all cross validation approaches tend to be conservative,} \bnote{We should discuss this and the sentence in abstract at our next meeting. They are conservative only in terms of selecting the model that has the best preditive performance; they do not miss selecting variables in the true support}
The linear predictor approach offers the best combination of performance and stability, and we recommend using it for regularization parameter selection in penalized Cox models.  We conclude by applying both proposed and existing approaches to two high dimensional data sets from real studies with time-to-event outcomes.

@@

The fundamental challenge of conducting cross validation for Cox regression is that the baseline hazard is not estimated from the model. Hence, another alternative approach to cross-validation would be to obtain an estimate $\hat{\Lambda}_{0}$ for the baseline hazard and use this in cross-validation.  There are several potential approaches to estimating $\Lambda_0$; in this manuscript, we chose to obtain the cross-validated linear predictors $\hat{\eta}^{cv}$ as in \eqref{eq:cv-lp}, then estimate $\Lambda_0$ via Kalbfleisch and Prentice's method \citep{Kalbfleisch2011}, treating these linear predictors as fixed inputs.\note{Maybe this isn't critical, but I still think we could just use $\hat{\Lambda}_0$ from the full data fit \ldots that estimator just makes more sense to me}\bnote{They do act a bit different in simulations. The baseline built from the full data performed worse. Discuss this next meeting}

@@

Simulation studies were conducted to compare how those methods behave relative to each other in selecting the tuning parameter $\lambda$ and the corresponding model for LASSO penalized regression. Survival times were generated from exponential distribution $h(t) = h_{0} \exp(X\beta)$, where $h_{0}$ is the prespecified baseline\note{huh?  so what is $h_0$?} \bnote{$h_0$ is 1 in all simulations. Basically, I first generated $X$, then the linear predictor $X\beta$; Then I compute the rate parameter from $h_0exp(X\beta)$ and inserted this to the \texttt{rexp} function to generate the survival times}; the covariate matrix $X$ with $n$ observations and $p$ covariates is generated from independent $\text{Normal}(0, 1)$ samples; and the coefficient vector $\beta$ is sparse, with most entries equal to zero. Censoring status indicators were drawn independently from the binomial distribution. All simulations were implemented in R \citep{R}. We varied the sample size, dimension, censoring percentage and signal strength of the simulated data sets to examine how these factors affect the performance of cross-validation.

@@

Estimation accuracy was measured using the mean squared error of the coefficients ($\MSE$). Suppose $\hat{\beta}(\lambda)$ is the set of coefficients estimated by a cox model with LASSO penalty at $\lambda$. Mean squared error $MSE$ is defined to be 
\bnote{
\begin{equation}
E\| \hat{\beta}(\lambda) - \beta \| ^2.
\end{equation}
}
\note{this is the equation for a scalar MSE, not a vector}

@@

As in Figure~\ref{Fig:mse-brier-c}, the standard and linear predictor approaches outperform the deviance residual and $V\&VH$ approaches.  The improvement is particularly noticeable in the most difficult setting: high-dimensional data with weak signal.  Table~\ref{Tab:sim} also provides some insight into why the deviance residual and $V\&VH$ approaches perform poorly relative to standard CV and the linear predictor approach: they act conservatively, consistently choosing larger $\lam$ values and selecting fewer variables. \note{if we're going to say something about ``selecting fewer variables'', should Table 1 include the average model size?} \bnote{I'm afraid that the average model size would be confusing to the readers. The most conversative model would have the smallest model size, and would be closest to the true model size}

@@

%\begin{figure}[h]
%    \centering
%		\includegraphics[height= 15cm ]{./figures/figure_auc_extended.png}
%    \caption{Cross-Validated AUC}
%\end{figure}
